### 2018-02-13 manuscript.tex
a bit more writing!

\TODO{resut intro here}

* experiment goal

* design

* source

* express the idea that this is just simple example that could be scaled up

First of all, we load the necessary libraries (see availability section for installation instructions):
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlkwd{library}\hlstd{(damr)}      \hlcom{# input DAM2 data}
\hlkwd{library}\hlstd{(zeitgebr)}  \hlcom{# periodogram computation}
\hlkwd{library}\hlstd{(sleepr)}    \hlcom{# sleep analysis}
\hlkwd{library}\hlstd{(ggetho)}    \hlcom{# behaviour visualisation}
\end{alltt}
\end{kframe}
\end{knitrout}

@@

\TODO{some concluding remarks}

* high scalability

* reproducibility

* from raw data to publication-quality figures

* flexibility

@@

@@

Markus, Patrick Kr{\"a}tschmer Rob

>>>>

In order to illustrate the workflow of \texttt{rethomics}, we decided to provide an example that is simple and reproducible, but which could be readily modified for more complicated cases.
DAM2 being widely adopted, and fruit flies being a model of choice for high throughput studies,
we decided to show how \texttt{rethomics} can be used to study circadian rhythm in fruit flies, using this type data.

We gathered a subset of the data from a recent publication \citationneeded{}, which was gracefully made publicly available by the authors\citationneeded{zenodo}.
It shows that flies, which normally become arrhythmic in a Light-Light (LL) regime, are rhythmic (with an period greater than 24h\TODO{phrasing}) when the gene NKCCox is expressed in clock neurons.
The authors have recorded the activity of both TIM/NKCCOK (NKCCox is expressed in their clock neurons) and TIM/+ (the appropriate control) in LL for six days.
Prior to LL, the animals were recorded for a baseline three or four days in normal LD, 12h:12h, conditions.
Two repetitions were performed resulting in a total of 58 animals.
The \texttt{metadata.csv} file, which describes exhaustively all individuals, as well as all the associated result files can be downloaded on zenodo\citationneeded{ibid}.

As a prerequisite, we downloaded the data, extracted the zip archive in our working directory.
We start the analysis by loading the necessary \texttt{rethomics} packages (see availability section for installation instructions):
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlkwd{library}\hlstd{(damr)}      \hlcom{# input DAM2 data}
\hlkwd{library}\hlstd{(zeitgebr)}  \hlcom{# periodogram computation}
\hlkwd{library}\hlstd{(sleepr)}    \hlcom{# sleep analysis}
\hlkwd{library}\hlstd{(ggetho)}    \hlcom{# behaviour visualisation}
\end{alltt}
\end{kframe}
\end{knitrout}

@@

This examples shows how \texttt{rethomics} can be used from the raw data to making publication quality figures and statistics.
We were able to comprehensively analyse classical circadian data with a few line of code.
Since the interface is programmatic, it adapts really well to much larger data sets and provides a large degree of freedom and flexibility.

@@

* How we have already used rethomics in two impactful papers

@@

Marcus Gosh, Patrick Kr{\"a}tschmer Rob

### 2018-02-13 results.Rnw
todolist
a bit more writing!

TODO description of the dataset here

* experiment goal

* design

* source

* express the idea that shis is just simple example that could be scaled up

First of all, we load the necessary libraries (see availability section for installation instructions):
<<message=FALSE>>=
library(damr)      # input DAM2 data
library(zeitgebr)  # periodogram computation
library(sleepr)    # sleep analysis
library(ggetho)    # behaviour visualisation
@

@@

Some conclusion here, TODO:

* high scalability

* reproducibility

* from raw data to publication-quality figures

* flexibility

>>>>

In order to illustrate the workflow of \texttt{rethomics}, we decided to provide an example that is simple and reproducible, but which could be readily modified for more complicated cases.
DAM2 being widely adopted, and fruit flies being a model of choice for high throughput studies,
we decided to show how \texttt{rethomics} can be used to study circadian rhythm in fruit flies, using this type data.

We gathered a subset of the data from a recent publication \citationneeded{}, which was gracefully made publicly available by the authors\citationneeded{zenodo}.
It shows that flies, which normally become arrhythmic in a Light-Light (LL) regime, are rhythmic (with an period greater than 24h\TODO{phrasing}) when the gene NKCCox is expressed in clock neurons.
The authors have recorded the activity of both TIM/NKCCOK (NKCCox is expressed in their clock neurons) and TIM/+ (the appropriate control) in LL for six days.
Prior to LL, the animals were recorded for a baseline three or four days in normal LD, 12h:12h, conditions.
Two repetitions were performed resulting in a total of 58 animals.
The \texttt{metadata.csv} file, which describes exhaustively all individuals, as well as all the associated result files can be downloaded on zenodo\citationneeded{ibid}.

As a prerequisite, we downloaded the data, extracted the zip archive in our working directory.
We start the analysis by loading the necessary \texttt{rethomics} packages (see availability section for installation instructions):
<<message=FALSE>>=
library(damr)      # input DAM2 data
library(zeitgebr)  # periodogram computation
library(sleepr)    # sleep analysis
library(ggetho)    # behaviour visualisation
@

@@

This examples shows how \texttt{rethomics} can be used from the raw data to making publication quality figures and statistics.
We were able to comprehensively analyse classical circadian data with a few line of code.
Since the interface is programmatic, it adapts really well to much larger data sets and provides a large degree of freedom and flexibility.
